End-to-End Verification of Initial and Transition Properties of GR(1) Designs in SPARK\thanks{Supported by AFRL contract FA8650-16-C-2642 and AFOSR grant RQCOR20-35.}

### 2020-01-28 doc/2020-CAV/paper.tex
Edits to free up space

\noindent 
In this context, the temporal operators $\always$ ``always,'' $\eventually$ ``eventually,'' and $\lnext$  ``next'' have the following meanings. 
A formula of the form $\always b$ holds if $b$ is true at every time step, 
$\eventually b$ holds if $b$ is eventually true at some future time step, 
$\always \eventually b$ holds if $b$ is true infinitely often in the future, 
and $\lnext b$ holds if $b$ is true at the next time step. 
It is assumed that at each time step, the environment chooses an input from $\inputs$, then the system chooses an output from $\outputs$ in response. 
The result of synthesis is a \emph{control protocol} or \emph{strategy} that can be expressed as a Moore machine. 
In a Moore machine, each state produces a unique set of system output values, 
and in the case of GR(1), synthesized Moore machines have the additional property that 
there is a unique set of environment input values that brings the system into each state.

Several tools for GR(1) synthesis are available. 
For example, RATSY %(Requirements Analysis Tool with Synthesis) 
\cite{bloem2010ratsy} has a focus on circuit design and can synthesize designs 
encoded in BLIF, %(Berkeley Logic Interchange Format),
Verilog, and HIF. %(HDL Intermediate Format). 
Similarly, Anzu \cite{jobstmann2007anzu} produces circuit designs in Verilog.
LTLMoP 
%(Linear Temporal Logic MissiOn Planning)  
\cite{finucane2010ltlmop} is focused on control of robots modeled as hybrid systems, 
and it synthesizes designs as hybrid controllers with handler modules to help connect controllers to simulated or real-world systems. 
TuLiP 
%(Temporal Logic Planning) 
\cite{TuLiP2011} has a similar focus and can synthesize controller implementations in Python. 
Slugs 
%(SmalL bUt Complete GROne Synthesizer) 
\cite{Ehlers2016} is architected to allow users to easily tailor synthesis algorithms for specific applications,
 e.g. to optimize criteria such as quick response, cost-optimality, and error-resilience, and it produces mathematical representations of controller designs.
Salty \cite{elliott2019salty} provides a front-end to Slugs that makes specifications easier to write and debug and 
a back-end that turns controller designs into executable software implementations in a variety of programming languages.

Though these tools implement synthesis algorithms that are theoretically ``correct-by-construction,'' 
tool implementation errors could still result in errors in synthesized controllers. 
We are therefore interested in ``end-to-end'' verification of synthesized controllers with respect to their original GR(1) specifications. 
Toward this end, we have extended Salty to produce software implementations of synthesized controller designs in SPARK. 
SPARK is both a programming language with a specification language and associated verification toolset \cite{hoang2015spark}. 
Though SPARK aims to perform fully automated verification, often the user must help guide the underlying provers by
annotating the code, e.g. with assertions and loop invariants. 
However, since controller designs synthesized from GR(1) specifications follow a regular structure, it was our hope that we could 
synthesize both the controller logic and the annotations necessary to automatically prove that controllers meet their functional specifications in SPARK. 
To date, we have achieved this goal for a subset of properties comprising GR(1) specifications, 
i.e. system initial and transition properties, for moderately-sized controllers.
In what follows, in \sectref{sec:implementation} we discuss the SPARK implementation. 
In \sectref{sec:caseStudies}, we demonstrate the efficacy of our approach on examples pulled from various sources.
In \sectref{sec:lessonsLearned}, we discuss lessons learned, including how to structure synthesized controllers for proof, 
examples in which SPARK revealed unexpected controller behaviors, and caveats related to the interpretation of GR(1) specifications. 
We end with concluding remarks in \sectref{sec:conclusions}, including a discussion of future work.

@@

To explain the structure of the controllers we generate in SPARK, consider a simple example of a controller for a traffic light. 
Color changes are triggered in every state in which an input signal's value is true.
Let this input be called $tick$.
The traffic light's color changes from red to green to yellow to red $\ldots$, repeating the cycle infinitely. 
Let the output variables then be $red$, $yellow$, and $green$.
The environment specifications are $\varphi^e_i = \top$, $\varphi^e_t = \top$, $\varphi^e_l = \always \eventually tick$, i.e.  
$tick$ can be true or false in the initial state and has no constraints on how it transitions from state to state, but it must be true infinitely often.
The system initial and liveness specifications are $\varphi^s_i = red \land \lnot yellow \land \lnot green$ and $\varphi^s_l =  \always \eventually green$, i.e.
the light starts as red but should infinitely often be green. The system transition specification is
%{\footnotesize
%\begin{align}
%\always & (\lnext ( (red \land \lnot yellow \land \lnot green) \lor (\lnot red \land yellow \land \lnot green) \lor \mbox{} \label{eq:mutex1} \\
%	& \quad \quad \quad (\lnot red \land \lnot yellow \land green)) \land \mbox{}  \label{eq:mutex2} \\
%	& \;\; (red \land \lnext tick \limplies \lnext green) \land (green \land \lnext tick \limplies \lnext yellow) \land \mbox{} \label{eq:change1} \\
%	& \quad \quad \quad (yellow \land \lnext tick \limplies \lnext red) \land \mbox{}  \label{eq:change2} \\
%	& \;\; (red \land \lnext \lnot tick \limplies red) \land (green \land \lnext \lnot tick \limplies green) \land \mbox{} \label{eq:same1} \\
%	& \quad \quad \quad (yellow \land \lnext \lnot tick \limplies yellow) \label{eq:same2})
%\end{align}
%}%
%That is, \eqnref{eq:mutex1} - \eqnref{eq:mutex2} express that the light should only be one color at a time; 
%\eqnref{eq:change1} - \eqnref{eq:change2} that the color should change from red to green to yellow and back to red whenever $tick$ is true; 
%and the color should remain the same whenever $tick$ is false.

@@

@@

\sloppy{
In SPARK, subunits consist of a specification and a body. 
For each synthesized controller, Salty automatically generates both a specification and body for a package named according to the Salty specification. 
The package specification for this traffic light example is shown in \figref{fig:trafficLightSpec}.
Type \lstinline{Controller} encodes a Moore machine suitable for representing the state of a synthesized controller. 
In the public part of the specification, \lstinline{Controller} (line 2) is declared as a private type so that the user cannot arbitrarily manipulate its state 
(possibly violating the controller's GR(1) specification). 
In the private part of the specification, \lstinline{Controller} (lines 33-37) is a record that stores the controller's internal state number, 
current input value(s), and the current output value(s). 
Its initial state is always the last possible state number. 
Each input and output is of type \lstinline{Boolean}.
When there are multiple inputs or outputs, they are wrapped in a record of type \lstinline{Environment} or \lstinline{System} (lines 4-6), respectively. 
Here, there is only one input, so it is not wrapped in a record. 
Since \lstinline{Controller} is a Moore machine, for each internal state, 
there is exactly one set of output values produced in that state. 
Furthermore, controllers synthesized from GR(1) specifications have the additional property that 
there is exactly one set of input values that brought the controller into that state. 
Therefore, \lstinline{Controller} includes a type invariant (lines 36-37) that captures this property, 
where functions \lstinline{State_To_Input_Mapping} and \lstinline{State_To_Output_Mapping} state whether the controller has the corresponding state number and input or output value(s), respectively.
They use lookup tables, \lstinline{State_To_Input_Equivalence} and \lstinline{State_To_Output_Equivalence}, defined in the package body, to do so.
Both of the functions used as invariants are declared (lines 26-29) with aspect \lstinline{Ghost}, indicating that they are intended mainly for proof purposes, 
i.e. they will create verification conditions related to type \lstinline{Controller} but will not be executed in the actual program, 
unless it is specified with a certain switch during compilation that they should be.
Executing ghost code is often used for debugging, e.g. to check whether an unproved property is actually true. 
In this case, removing the ghost code can save significant memory, 
especially on bigger examples where the tables encoded for functions  \lstinline{State_To_Input_Mapping} and \lstinline{State_To_Output_Mapping} 
would have to store information on thousands of states.
The logic for these functions is given in the body (not shown here), but for instance in state 1, \lstinline{State_To_Input_Equivalence} returns \lstinline{False}, 
since \lstinline{tick = False} as the system enters state 1, and \lstinline{State_To_Output_Equivalence} returns \lstinline{System'(red => True, yellow => False, green => False)}, 
since $red$ is True and $yellow$ and $green$ are false as the system enters state 1.
 }

The public function \lstinline{Is_Init} (line 8) checks whether the controller is in its initial state, i.e. no inputs have yet been received. 
The public function \lstinline{Env_Init} (line 9) checks whether input(s) satisfy $\varphi_e^i$. 
It is implemented as an expression function, i.e. the implementation is given directly in the specification and not the body, 
because all terms needed to define it are visible in the public part of the specification. 
In this example, since $\varphi_e^i = \top$ as pulled from the Salty specification, this function always simply returns \lstinline{True}. 
The public function \lstinline{Sys_Init} (lines 10-11) checks whether outputs(s) satisfy $\varphi_s^i$. 
It is implemented as an expression function for the same reason.  
As pulled from the Salty specification, $\varphi^s_i = red \land \lnot yellow \land \lnot green$, so this function returns the value of
the expression \lstinline{S.red and not S.yellow and not S.green}.
But unlike \lstinline{Is_Init} and \lstinline{Env_Init}, it is marked with aspect \lstinline{Ghost} because it is mainly used for proof and does not need to be executed. 
\lstinline{Is_Init} and \lstinline{Env_Init} are used for proof but are also callable in functional code. 
We chose to make these non-ghost functions for reasons related to the meaning of GR(1) specifications. 
Recall that GR(1) specifications have the form $\varphi_e \limplies \varphi_s$. 
If $\varphi_e$ ever becomes false, i.e. if the environment produces input value(s) that violate $\varphi_e$, then the specification as a whole 
is satisfied regardless of whether the system produces output value(s) that satisfy $\varphi_s$. 
In theory, the system in this case could produce arbitrary outputs and still satisfy the overall specification. 
In practice, we believe a user would generally want to know that the environment violated its specification, 
so that the user could either choose the system output value(s) explicitly or fall back to some other error handling routine. 
Therefore, a user needs to be able to check inputs with \lstinline{Env_Init} if \lstinline{Is_Init} returns true, 
which is why both are callable. At the moment, they are not used in the code of the \lstinline{Move} procedure.
Public functions \lstinline{Env_Trans} and \lstinline{Sys_Trans} (lines 13-16) check whether the next set of input value(s) and output value(s) 
satisfy $\varphi_e^t$ and $\varphi_s^t$, respectively. 
For the same reasons as above, \lstinline{Sys_Trans} is a ghost function but \lstinline{Env_Trans} is not.
Note that \lstinline{Env_Trans} has a precondition that the controller must not be in its initial state, 
since $\varphi_e^t$ can depend on both the current and next set of input value(s). 
This precondition is also necessary for \lstinline{Sys_Trans}, because in the initial state, the attributes of the \lstinline{Controller} are not initialized, and the \lstinline{Sys_Trans} function might read these values. 
As with \lstinline{Env_Init} and \lstinline{Sys_Init}, the logic for these functions is pulled from the Salty specification and 
implemented in the body (not shown), since they make use of input and output values stored in the \lstinline{Controller}, whose fields are private.

@@

The public procedure \lstinline{Move} (lines 18-23) transitions a controller based on its current state and 
next set of input value(s), and it produces the next set of output value(s). 
It has a precondition that if the controller is in its initial state, 
inputs must satisfy $\varphi_e^i$; otherwise they must satisfy $\varphi_e^t$.
The aspect \lstinline{Contract_Cases} specifies additional sets of preconditions paired with postconditions, 
where the set of all preconditions must be mutually exclusive and exhaustively cover the entire input space.
The \lstinline{others} keyword can be used to cover the set of all input conditions not covered in any explicit cases of the contract. 
Note that for the left-hand side of each case (left of the \lstinline{=>}), variable names refer to variable values before evaluation of the subprogram; 
for the right-hand side, they refer to variable values after evaluation.
For the right-hand side, the aspect \lstinline{Old} can therefore be used to reference the value of a variable before evaluation of the subprogram. 
Combined with the previous precondition, \lstinline{Contract_Cases} asserts that if the controller is in its initial state, then after execution of \lstinline{Move}, 
the first set of output value(s) produced should satisfy $\varphi_s^i$ and the controller should no longer be in its initial state; 
if it is not in its initial state, then the output value(s) produced should satisfy $\varphi_s^t$, 
evaluated based on the most recent input and output values stored in \lstinline{C'Old}, the next input value(s) just provided (in this case in \lstinline{tick}), 
and the next output values just generated (in this case in the record \lstinline{System}). 
This set of contract cases embodies our main proof goal, i.e. verification of system initial and transition properties from the original GR(1) specification for the controller.

A fragment of the body of \lstinline{Move} is shown in \figref{fig:moveBody}. 
Note that there are cases that can lead to \lstinline{Program_Error}. 
This is because case statements require all possible cases to be covered, so 
we programmatically use \lstinline{others} to cover all possible input combinations 
that would not be allowed due to $\varphi_e^i$ or $\varphi_e^t$. 
In the traffic light example, they are unnecessary because all possible combinations of input values are 
allowed out of each state. 
In any case, SPARK will prove that these cases are not reachable.

>>>>

\noindent 
In this context, the temporal operators $\always$ ``always,'' $\eventually$ ``eventually,'' and $\lnext$  ``next'' have the following meanings. 
A formula of the form $\always b$ holds if $b$ is true at every time step, 
$\eventually b$ holds if $b$ is eventually true at some future time step, 
$\always \eventually b$ holds if $b$ is true infinitely often in the future, 
and $\lnext b$ holds if $b$ is true at the next time step. 
It is assumed that at each time step, the environment chooses an input from $\inputs$, then the system chooses an output from $\outputs$ in response. 
The result of synthesis is a \emph{control protocol} or \emph{strategy} that can be expressed as a Moore machine. 
%Note that in the case of GR(1), synthesized Moore machines not only produce a single set of system output values in each state, 
%there is a unique set of environment input values that brings the controller into each state.

Several tools for GR(1) synthesis are available. 
For example, RATSY %(Requirements Analysis Tool with Synthesis) 
\cite{bloem2010ratsy} has a focus on circuit design and synthesizes designs 
encoded in BLIF, %(Berkeley Logic Interchange Format),
Verilog, and HIF. %(HDL Intermediate Format). 
Similarly, Anzu \cite{jobstmann2007anzu} produces circuit designs in Verilog.
LTLMoP 
%(Linear Temporal Logic MissiOn Planning)  
\cite{finucane2010ltlmop} is focused on control of robots modeled as hybrid systems, 
and it synthesizes designs as hybrid controllers with handler modules to help connect controllers to simulated or real-world systems. 
TuLiP 
%(Temporal Logic Planning) 
\cite{TuLiP2011} has a similar focus and can synthesize controller implementations in Python. 
Slugs 
%(SmalL bUt Complete GROne Synthesizer) 
\cite{Ehlers2016} is architected to allow users to tailor synthesis algorithms,
 e.g. to optimize criteria such as quick response, cost-optimality, and error-resilience, and it produces mathematical representations of controller designs.
Salty \cite{elliott2019salty} provides a front-end to Slugs that makes specifications easier to write and debug and 
a back-end that turns controller designs into executable software implementations in a variety of programming languages.

Though these tools implement synthesis algorithms that are theoretically ``correct-by-construction,'' 
tool implementation errors could still result in errors in synthesized controllers. 
We are therefore interested in ``end-to-end'' verification of synthesized controllers. % with respect to their original GR(1) specifications. 
Toward this end, we have extended Salty to produce software implementations of synthesized controller designs in SPARK. 
SPARK is both a programming language with a specification language and associated verification toolset \cite{hoang2015spark}. 
Though SPARK aims to perform fully automated verification, often the user must help guide the underlying provers by
annotating the code, e.g. with assertions and loop invariants. 
However, since controller designs synthesized from GR(1) specifications follow a regular structure, it was our hope that we could 
synthesize both the controller logic and the annotations necessary to automatically prove that controllers meet their GR(1) specifications in SPARK. 
To date, we have achieved this goal for a subset of properties comprising GR(1) specifications, 
i.e. system initial and transition properties, for moderately-sized controllers.
In what follows, in \sectref{sec:implementation} we discuss the SPARK implementation. 
In \sectref{sec:caseStudies}, we discuss our results on examples pulled from various sources.
In \sectref{sec:lessonsLearned}, we discuss lessons learned, including how to structure synthesized controllers for proof, 
examples in which SPARK revealed unexpected controller behaviors, and caveats related to GR(1) specifications. 
We end with concluding remarks in \sectref{sec:conclusions}, including a discussion of future work.

@@

To explain the structure of our synthesized SPARK controllers, consider a simple example of a traffic light controller. 
Let the controller's single input be $tick$. 
Color changes occur in every state in which $tick$ is true.
The traffic light's color cycles infinitely over the sequence red $\rightarrow$ green $\rightarrow$ yellow $\rightarrow$ red $\rightarrow$ $\ldots$.
Let the output variables then be $red$, $yellow$, and $green$.
The environment specifications are $\varphi^e_i = \top$, $\varphi^e_t = \top$, $\varphi^e_l = \always \eventually tick$, i.e.  
$tick$ can be true or false in the initial state and has no constraints on how it transitions from state to state, but it must be true infinitely often.
The system initial and liveness specifications are $\varphi^s_i = red \land \lnot yellow \land \lnot green$ and $\varphi^s_l =  \always \eventually green$, i.e.
the light starts as red but should infinitely often be green. The system transition specification is
%{\footnotesize
%\begin{align}
%\always & (\lnext ( (red \land \lnot yellow \land \lnot green) \lor (\lnot red \land yellow \land \lnot green) \lor \mbox{} \label{eq:mutex1} \\
%	& \quad \quad \quad (\lnot red \land \lnot yellow \land green)) \land \mbox{}  \label{eq:mutex2} \\
%	& \;\; (red \land \lnext tick \limplies \lnext green) \land (green \land \lnext tick \limplies \lnext yellow) \land \mbox{} \label{eq:change1} \\
%	& \quad \quad \quad (yellow \land \lnext tick \limplies \lnext red) \land \mbox{}  \label{eq:change2} \\
%	& \;\; (red \land \lnext \lnot tick \limplies red) \land (green \land \lnext \lnot tick \limplies green) \land \mbox{} \label{eq:same1} \\
%	& \quad \quad \quad (yellow \land \lnext \lnot tick \limplies yellow) \label{eq:same2})
%\end{align}
%}%
%That is, \eqnref{eq:mutex1} - \eqnref{eq:mutex2} express that the light should only be one color at a time; 
%\eqnref{eq:change1} - \eqnref{eq:change2} that the color should change from red to green to yellow and back to red whenever $tick$ is true; 
%and the color should remain the same whenever $tick$ is false.

@@

\renewcommand{\floatpagefraction}{.8}%

@@

In SPARK, subunits consist of a specification and a body, 
which Salty generates. 
% For each synthesized controller, Salty automatically generates both a specification and body for a package named according to the Salty specification. 
The package specification for this traffic light example is shown in \figref{fig:trafficLightSpec}.
Type \lstinline{Controller} encodes a Moore machine representing the synthesized controller. 
In the public part of the specification, \lstinline{Controller} (line 2) is declared as a private type so that the user cannot arbitrarily manipulate its state.  
% (possibly violating the controller's GR(1) specification). 
In the private part of the specification, \lstinline{Controller} (lines 33-37) is a record that stores the controller's internal state number, 
current input value(s), and the current output value(s). 
Its initial state is always the last possible state number. 
Each input and output is of type \lstinline{Boolean}. 
When there are multiple inputs or outputs, they are wrapped in a record of type \lstinline{Environment} or \lstinline{System} (lines 4-6), respectively. 
Here, there is only one input, so it is not wrapped in a record. 
Since \lstinline{Controller} is a Moore machine, for each internal state, 
there is exactly one set of output values produced in that state. 
Furthermore, controllers synthesized from GR(1) specifications have the additional property that 
there is exactly one set of input values that brought the controller into that state. 
Therefore, \lstinline{Controller} includes a type invariant (lines 36-37) that captures this property, 
where functions \lstinline{State_To_Input_Mapping} and \lstinline{State_To_Output_Mapping} state whether the controller has the corresponding state number and input or output value(s), respectively. 
% They use lookup tables, \lstinline{State_To_Input_Equivalence} and \lstinline{State_To_Output_Equivalence}, defined in the package body, to do so.
Both of these functions are declared (lines 26-29) with aspect \lstinline{Ghost}, indicating that they are intended mainly for proof purposes, 
i.e. they will create verification conditions related to type \lstinline{Controller} but will not be executed in the actual program, 
unless it is specified with a certain switch during compilation that they should be.
Executing ghost code is often used for debugging, e.g. to check whether an unproved property is actually true. 
In this case, removing the ghost code can save significant memory, 
since \lstinline{State_To_Input_Mapping} and \lstinline{State_To_Output_Mapping} 
internally encode large lookup tables that have to store information on every state of the controller, 
which is often thousands of states or even millions of states for the largest example in our database.
The logic for these functions is given in the body (not shown here), but for instance when \lstinline{C.State = 1}, 
\lstinline{State_To_Input_Mapping} returns \lstinline{True} if and only if \lstinline{C.tick = False} and 
\lstinline{State_To_Input_Mapping} returns \lstinline{True} if and only if 
\sloppy{\lstinline{C.S = System'(red => True, yellow => False, green => False)}}, 
since \lstinline{tick = False} brings the system into state 1 and $red$ is true and $yellow$ and $green$ are false in state 1.

The public function \lstinline{Is_Init} (line 8) checks whether the controller is in its initial state, i.e. no inputs have yet been received. 
The public function \lstinline{Env_Init} (line 9) checks whether input(s) satisfy $\varphi_e^i$. 
It is implemented as an expression function, i.e. the implementation is given directly in the specification, 
because all terms needed to define it are visible in the public part of the specification. 
In this example, since $\varphi_e^i = \top$ as pulled from the Salty specification, this function always simply returns \lstinline{True}. 
The public function \lstinline{Sys_Init} (lines 10-11) checks whether outputs(s) satisfy $\varphi_s^i$. 
It is implemented as an expression function for the same reason.  
As pulled from the Salty specification, $\varphi^s_i = red \land \lnot yellow \land \lnot green$, so this function returns the value of
the expression \lstinline{S.red and not S.yellow and not S.green}.
But unlike \lstinline{Is_Init} and \lstinline{Env_Init}, it is marked with aspect \lstinline{Ghost} because it is mainly used for proof and does not need to be executed. 
\lstinline{Is_Init} and \lstinline{Env_Init} are used for proof but are also callable in functional code. 
We chose to make these non-ghost functions for reasons related to the meaning of GR(1) specifications. 
Recall that GR(1) specifications have the form $\varphi_e \limplies \varphi_s$. 
If $\varphi_e$ ever becomes false, i.e. if the environment produces input value(s) that violate $\varphi_e$, then the specification as a whole 
is satisfied regardless of whether the system produces output value(s) that satisfy $\varphi_s$. 
In theory, the system in this case could produce arbitrary outputs and still satisfy the overall specification. 
In practice, we believe a user would generally want to know that the environment violated its specification, 
so that the user could either choose the system output value(s) explicitly or fall back to some other error handling routine. 
Therefore, a user needs to be able to check inputs with \lstinline{Env_Init} if \lstinline{Is_Init} returns true, 
which is why both are callable. At the moment, they are not used in the code of the \lstinline{Move} procedure.
Public functions \lstinline{Env_Trans} and \lstinline{Sys_Trans} (lines 13-16) check whether the next set of input value(s) and output value(s) 
satisfy $\varphi_e^t$ and $\varphi_s^t$, respectively. 
For the same reasons as above, \lstinline{Sys_Trans} is a ghost function but \lstinline{Env_Trans} is not.
Note that \lstinline{Env_Trans} has a precondition that the controller must not be in its initial state, 
since $\varphi_e^t$ can depend on both the current and next set of input value(s). 
This precondition is also necessary for \lstinline{Sys_Trans}, because in the initial state, the attributes of the \lstinline{Controller} are not initialized, and the \lstinline{Sys_Trans} function might read these values. 
As with \lstinline{Env_Init} and \lstinline{Sys_Init}, the logic for these functions is pulled from the Salty specification and 
implemented in the body (not shown), since they make use of input and output values stored in the \lstinline{Controller}, whose fields are private.

@@

The public procedure \lstinline{Move} (lines 18-23) transitions a controller based on its current state and 
next set of input value(s), and it produces the next set of output value(s). 
It has a precondition that if the controller is in its initial state, 
inputs must satisfy $\varphi_e^i$; otherwise they must satisfy $\varphi_e^t$.
The aspect \lstinline{Contract_Cases} specifies additional sets of preconditions paired with postconditions, 
where the set of all preconditions must be mutually exclusive and exhaustively cover the entire input space.
The \lstinline{others} keyword can be used to cover the set of all input conditions not covered in any explicit cases of the contract. 
Note that for the left-hand side of each case (left of the \lstinline{=>}), variable names refer to values before evaluation of the subprogram; 
for the right-hand side, they refer to values after evaluation.
Therefore, the aspect \lstinline{Old} can be used on the right to reference the value of a variable before evaluation of the subprogram. 
Combined with the previous precondition, \lstinline{Contract_Cases} asserts that if the controller is in its initial state, then after execution of \lstinline{Move}, 
the first set of output value(s) produced should satisfy $\varphi_s^i$ and the controller should no longer be in its initial state; 
if it is not in its initial state, then the output value(s) produced should satisfy $\varphi_s^t$, 
evaluated based on the most recent input and output values stored in \lstinline{C'Old}, the next input value(s) just provided (in this case in \lstinline{tick}), 
and the next output values just generated (in this case in the record \lstinline{System}). 
This set of contract cases embodies our main proof goal, i.e. verification of system initial and transition properties from the original GR(1) specification for the controller.

A fragment of the body of \lstinline{Move} is shown in \figref{fig:moveBody}. 
Note that there are cases that can lead to \lstinline{Program_Error}. 
This is because case statements require all possible cases to be covered, so 
we programmatically use \lstinline{others} to cover all possible input combinations 
that would not be allowed due to $\varphi_e^i$ or $\varphi_e^t$. 
In the traffic light example, these are unnecessary because all possible combinations of input values are 
allowed out of each state. 
In any case, SPARK will prove that such cases are not reachable.
